### 2024-09-27 paper/paper.qmd
Drafted introduction and data sections of paper

---
title: "Emergency Medical Response in Toronto: An Analysis of Ambulance Dispatches"
# subtitle: "My subtitle if needed"
author: 
  - Muhammad Abdullah Motasim
thanks: "Code and data are available at: https://github.com/abdullah-motasim/Analyzing-Ambulance-Response-Times"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
table-of-contents: true
toc-depth: 2
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| 
# Package loading
# install.packages("modeest")
library(tidyverse)
library(dplyr)
library(here)
library(tibble)
library(opendatatoronto)
library(ggplot2)
library(knitr)
library(modeest)

# Read data
raw_data <- read_csv(here("data/raw_data/raw_ambulance_response_data.csv"))
cleaned_data <- read_csv(here("data/analysis_data/cleaned_ambulance_response_data.csv"))
```

Quorto can only read data we load into it, so if we have data in global enviornment, quorto isnt able to use it

use (here::here("/path/to/folder)) to force quorto to look for folder outside of quorto working directory

use citation("library name") to get citation for a library

ggplot(cleaned_data, aes(x=year))+ geom_bar(position = "dodge")+ theme_minimal()

ggplot(cleaned_data, aes(x=year, fill=(units_arrived_at_scene)))+ geom_bar(position = "dodge")+ theme_minimal()+ \# theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Introduction

Efficient ambulances response times play a key role in saving saving peoples lives, especially in severe and sudden health complications such as cardiac arrests, trauma, and motor vehicle accidents. A medical study from 2011 found for cardiac arrest a 1 minute reduction in response time improved a persons odds of survival by 24% @citeBMJ. Furthermore, Toronto is facing issues retaining Emergency Medical Services (EMS) workers @citeCBCParamedicRetention leading to increased call wait times @citeCBC911WaitTimes and lower number of ambulances available to be dispatched @citeCBCAmbulanceResponseTimes.

This paper hopes to analyze the number and types of dispatch requests to EMS services within Toronto to identify the reason for the reduced amount of ambulances available and the increased call wait times. This is achieved utilizing a data set procured from OpenDataToronto @citeOpenDataTorontoWebsite containing information on the number of dispatches made per year, the severity of these reports and the number of units dispatched. We utilize this information to determine \<talk about how we found that the number of calls was increasing each year and the so was the severity. Analyze the severity data to see of higher severity means more dispatchers and if so say that the increase in the severity led to more calls and also more units needed per call, coupled with ems workers facing shortages and burnout led to poor ems services available to Canadians. Say thus is important as experts can now better tackle the underlying issues causing delayed ems response times, saving lives.\>

The remainder of this paper is structured as follows. @sec-data discusses the datatypes included in the raw data, the cleaning process for the data, and reason for selecting the data set we did. @sec-results analyzes the trends and correlations present between different variables utilizing tabular and graphical means. @sec-discussion the results of @sec-results going into detail on what these results can tell us about the cause for delayed ems response time. Lastly, @sec-nextsteps discusses limitations within the data and steps for improvement.

# Data {#sec-data}

## Raw Data

The data utilized for the analysis was sourced from the Open Data Toronto website and was read into the paper using the opendatatoronto library @citeOpenDataToronto, and all analysis was performed using R @citeR and the following packages: tidyverse @citeTidyverse, janitor @citeJanitor, lubridate @citeLubridate, truncnorm @citeTruncnorm, purrr @citePurrr, dplyr @citeDplyr, here @citeHere, ggplot2, @citeGgplot2, and knitr @citeKnitr. The chosen data set is called "Paramedic Services Incident Data" @citeOpenDataTorontoDataset and contains information about ambulances dispatches sent within Toronto from 2017-2022, the data is refreshed annually and was last updated October 5, 2023. The data is sourced and published by Toronto Paramedic Services meaning they decide the type of data collected and the method of collection. This set was chosen due to the large amount of data it contained, with over 1.7 million entries over the 5 years recorded meaning we could analyze trends over multiple years.

The raw downloaded data is shown in @tbl-raw-data-1 and @tbl-raw-data-2, as you can see it contains a total of 9 variables:

-   **ID** - Incident ID number associated with the Computer Aided Dispatch (CAD) record.
-   **Dispatch_Time** - Time the first Paramedic unit was assigned to the incident.
-   **Incident_Type** - Type of incident reported; possible values are:
    -   '-' (no value)
    -   Airport Standby
    -   Emergency Transfer
    -   Fire
    -   Medical
    -   Motor Vehicle Accident
-   **Priority_Number** - Priority level determined using the Medical Priority Dispatch System (MPDS), (see @sec-appendixA for details).
-   **Units_Arrived_At_Scene** - Number of Paramedic units that arrived on scene of incident
-   **Forward_Sortation_Area** - General location of incident based on Postal Code Forward Sortation Area

-   **Field Name** - Gives names of field such as ID, Dispatch_Time, Incident_Type, etc.
-   **Desctiption/Definition** - Describes what each field name means
-   **Comments/Examples** - Provides additional information on field name descriptions

```{r}
#| label: tbl-raw-data-1
#| tbl-cap: First 6 entires of the Paramedic Services Incident Dataset
#| echo: false
#| warning: false
#| message: false

head_raw_data <- head(raw_data)
# Note there is a type in Description within the original data set itself
knitr::kable(head_raw_data[,1:5], "simple",   col.names = c('ID', 'Dispatch_Time', 'Incident_Type', 'Priority_Number', 'Units_Arrived_At_Scene'))
```

```{r}
#| label: tbl-raw-data-2
#| tbl-cap: First 6 entires of the Paramedic Services Incident Dataset (cont.)
#| echo: false
#| warning: false
#| message: false
# Table is broken into 2 sections to show all data on a single page
knitr::kable(head_raw_data[,6:9], "simple",   col.names = c( 'Forward_Sortation_Area', 'Field Name', 'Desctiption/Definition', 'Comments/Examples'))
```

## Cleaned Data

During the creation of the raw data set the Field Names, Desctiption/Definition, and Comments/Examples columns were added as part of the process combining the data for all 5 years. However, we don't use these columns within our numerical analysis so they were removed within the data cleaning to reduce the data size. Furthermore, the dispatch_time column was formatted from type dbl to type POSIXct and two new columns were added containing the year and month of the report in order to allow for efficient plotting of dates. Lastly, the cleaned data was tested to ensure no missing values and all given values fell within their predetermined values. The cleaned data is shown in @tbl-cleaned-data-1 and @tbl-cleaned-data-2, @fig-year-vs-dispatches showcases the total number of dispatches per year.

```{r}
#| label: tbl-cleaned-data-1
#| tbl-cap: First 6 entires of the cleaned ambulance response data
#| echo: false
#| warning: false
#| message: false

head_cleaned_data <- head(cleaned_data)
# Note there is a type in Description within the original data set itself
knitr::kable(head_cleaned_data[,1:5], "simple",   col.names = c('id', 'dispatch_time', 'incident_type', 'priority_number', 'units_arrived_at_scene'))
```

```{r}
#| label: tbl-cleaned-data-2
#| tbl-cap: First 6 entires of the cleaned ambulance response data (cont.)
#| echo: false
#| warning: false
#| message: false

# Table is broken into 2 sections to show all data on a single page
knitr::kable(head_cleaned_data[,6:8], "simple",   col.names = c( 'forward_sortation_area', 'year', 'month'))
```

```{r}
#| label: fig-year-vs-dispatches
#| fig-cap: Year vs. Total Number of Dispatches
#| echo: false
#| warning: false
#| message: false

cleaned_data |>
ggplot(aes(x=year))+ 
  xlab("Year")+
  ylab("# of Dispatches")+
  geom_bar(position = "dodge")+ 
  scale_x_continuous(breaks = seq(min(cleaned_data$year), max(cleaned_data$year), by = 1)) + 
  theme_minimal()

```

## Summary of Data
We are only interested in the trend in dispatches over the years and @fig-year-vs-dispatches shows that there isn't much change in total number of dispatches over years meaning any calculated statistic on dispatch_time would not be of use. Thus,the only data types for which we can compute useful statistics are incident_type, priority_number, and units_arrived_at_scene. @tbl-cleaned-data-code shows the summary of these calculated statistics with mean for continuous data and mode for categorical data.

```{r}
#| label: tbl-cleaned-data-code
#| tbl-cap: Summary of Incident Data Statistics
#| echo: false
#| warning: false
#| message: false

# calculate statistics
mode_incident_type <- mlv(cleaned_data$incident_type, method = "mfv")
mode_priority_number <- mlv(cleaned_data$priority_number, method = "mfv")
mean_units_at_scene <- mean(cleaned_data$units_arrived_at_scene)

# Create a data frame to show values
summary_table <- data.frame(
  Field_Name = c("Units Arrived at Scene", "Mode Incident Type", "Mode Priority Number"),
  Value = c(mean_units_at_scene, mode_incident_type, mode_priority_number)
)

# Print the table
kable(summary_table)
```

# Results {#sec-results}

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

# Weaknesses and next steps {#sec-nextsteps}

Data doesn't have arrival times

\newpage

\appendix

# Appendix {.unnumbered}

## Appendix A {#sec-appendixA}

Information on priority levels of Medical Priority Dispatch System (MPDS) ratings was sourced from @citePriorityDocument. Below is the severity of each level.

| Priority \# |                                Description |
|-------------|-------------------------------------------:|
| 5           |                       Alpha (Low Priority) |
| 4           |                       Bravo (Mid Priority) |
| 3           |        Charlie (Possibly Life Threatening) |
| 1           |                   Delta (Life Threatening) |
| 9           |       Echo (Full Arrest or Imminent Death) |
| 11          |         Alpha1 (Most Urgency within Alpha) |
| 12          |          Alpha2 (Mid Urgency within Alpha) |
| 13          |        Alpha3 (Least Urgency within Alpha) |
| 14          | Code 2 (Non-emergency transport requests.) |

\newpage

# References